The Aquarius Stream Progenitor Was Not a Globular Cluster

### 2012-11-20 2012-aqs/draft-aqs.tex
Updates

% is it a real substructure

% MRS follow-up by de boer

@@

We seek to investigate the globular cluster origin claim made by \citet{de_Boer;et-al_2012}. We present a detailed chemical abundance analysis for six confirmed Aquarius stream members observed using the Magellan Inamori Kyocera Echelle spectrograph \citep{Bernstein;et-al_2002} on the Magellan telescope. Details of the observations and data reduction are outlined in the following section. The data analysis is presented in \S\ref{sec:analysis} and a detailed discussion of these results resides in \S\ref{sec:discussion}, as well as an alternative hypothesis for the origin of the Aquarius stream. In \S\ref{sec:conclusion} we present our conclusions and critical interpretations.

@@

\subsection{Radial Velocities}
\label{sec:radial-velocities}
The radial velocity for each star was determined in a two step method. Initially each normalised, stitched spectrum was cross-correlated with a synthetic spectrum of a K0 giant with $T_{eff} = 4500$\,K, $\log{g} = 1.5$, and [M/H] $= -1.0$ across the wavelength range $8450 - 8700$ \AA{}. Velocities found during our cross-correlation are typically within 1\,km s$^{-1}$, and this measurement is used to place each spectrum at its approximate rest wavelength. The rest spectrum was used to measure equivalent widths for N absorption lines in each spectrum (\S\ref{sec:line-measurements}). The measurement of equivalent widths involves fitting a Gaussian profile to each absorption feature. Thus, in addition to an equivalent width, the measured central wavelength is found from the Gaussian centroid. The ratio between the measured central wavelength and the laboratory rest wavelength is primarily determined by the stellar radial velocity. Thus, we have N measurements of the stellar radial velocity.

The second step of our radial velocity determination is to find the mean offset and standard deviation from this distribution of line velocity measurements. Figure \ref{fig:line-velocities} shows the line velocities for HD440077 after  being placed at rest using our cross-correlation velocity of X.XX\,km s$^{-1}$. As expected, the mean offset is small ($-1.1$\,km s$^{-1}$), and our standard deviation is X.XX\,km s$^{-1}$ from N line measurements. This provides us with a final measured radial velocity of $X.XX \pm X.XX$\,km s$^{-1}$. This process was repeated for all standards and observations. The radial velocities published in Table \ref{tab:observations} are the final values from this two-step method.

@@

\subsection{Model Atmospheres}
We employed the ATLAS9 stellar atmospheres of \citet{Castelli;Kurucz_2003} for this analysis. These one-dimensional models ignore any centre-to-limb spatial variations, assume local thermodynamic equilibrium as well as overall hydrostatic equilibrium, and assumes no convective overshoot from the photosphere. The stellar parameter spacing for these grid models is 250\,K, 0.5 dex in surface gravity, 0.5 dex in [M/H] and 0.1 dex in [$\alpha$/Fe]. We interpolated the stellar densities, temperatures, electron pressures at all stellar depths between atmospheres using the Quickhull algorithm during our analyses. Quickhull is reliant on Delaunay tessellation, which suffers from extremely skewed cells when the grid points vary in size by orders of magnitude -- as $T_{eff}$ values do compared to $\log{g}$ or [M/H]. If unaccounted for, these asymmetric cells can manifest as non-negligible errors in interpolated photospheric quantities across all photospheric depths. As such, we scaled each stellar parameter between zero and unity prior to interpolation. During this step, $\log{g}$, [Fe/H] and [$\alpha$/Fe] points were scaled such that we simultaneously interpolate linearly in $T_{eff}$ and logarithmically in all other parameters.

\subsection{Stellar Parameters}
The most recent version of the spectral synthesis code MOOG \citep{Sneden;et-al_1973} has been used to derive individual line abundances and stellar parameters. This version employs a more accurate treatment of Rayleigh scattering \citet{Sobeck;et-al_2011}, which is particularly important for transitions blueward of 450\,nm but less relevant for this analysis as most of our line measurements are redward of 450\,nm. During our analysis we employed the updated solar abundances of \citet{Asplund;et-al_2009}.

The effective temperature for each star was found by demanding a zero-trend in excitation potential and abundance for each measured Fe\,I line. In the same way, the microturbulence was found by demanding a zero trend in reduced equivalent width against abundance. Linear relationships with slopes of |0.03| dex were considered to be converged. Surface gravity was found by forcing the mean Fe\,I and Fe\,II abundances to be equal, whilst maintaining zero trends with the excitation potential, reduced equivalent width and abundance. After these parameters had converged, the model metallicity was exactly matched to that of our line abundances. Line abundances that were unusually deviant ($>3\sigma$) from the mean were removed.

Stellar parameters for our standard stars are listed in Table \ref{fig:standards-stellar-parameters}. We find an excellent agreement with that of \citet{Gratton;et-al} and \citet{Fulbright;et-al_1997} and no systematic differences are present. In contrast, the stellar parameters we deduce for our Aquarius stream program stars are significantly different both from those found by the RAVE pipeline \citep{}, as listed in \citet{Williams;et-al_2011}, and those found by \citet{Wylie-de-boer;et-al_2012} with medium-resolution spectroscopy.

@@

@@

@@

In contrast, we present 52 Fe\,I and 13 Fe\,II lines that have been cleanly selected with great

The NIST database contains X

>>>>

% is it a real substructure
\citet{Williams;et-al_2011} quantified the statistical significance of the Aquarius substructure with comparisons to the Besancon and Galaxia models. After populating the models, the galaxy representations were discretized in $\Delta{}V_{hel}$ and $\Delta{}l$ grid blocks.

There is no reason to suspect that the Aquarius stream is not a real substructure.

% What is it?
Given the extent of stellar debris the Sagittarius dwarf has littered in the Milky Way, it is reasonable to suspect the Aquarius stream may simply be Sagittarius in origin. Although the metallicities reported by \citet{Williams;et-al_2011} are not dissimilar from the stream, the distances and $V_{Z}$, $V_\phi$ for Aquarius are quite distinct from Sagittarius. \citet{Williams;et-al_2011} concluded that the newly discovered stream could not be positively associated with the Monoceros stream, Hercules-Aquila cloud, or either the Canis Major or Virgo Overdensities.

% MRS follow-up by de boer
In order to investigate the Aquarius progenitor and any possible association with existing substructures, \citet{Wylie-de_boer;et-al_2012} observed six Aquarius stream members with medium resolution (R = 25,000) spectroscopy. Their data indicates the stream is chemically coherant, with a dispersion of only 0.1\,dex. Based on sodium and nickel abundance ratios with respect to iron, they claim the Aquarius stream chemical signatures unambiguously represent those of a globular cluster. Additional abundances of oxygen, magnesium and aluminium are reported to solidify this claim.

@@

We seek to investigate the globular cluster origin claim made by \citet{wylie-de_Boer;et-al_2012}. We present a detailed chemical abundance analysis for six confirmed Aquarius stream members observed using the Magellan Inamori Kyocera Echelle spectrograph \citep{Bernstein;et-al_2002} on the Magellan telescope. Details of the observations and data reduction are outlined in the following section. The data analysis is presented in \S\ref{sec:analysis} and a detailed discussion of these results resides in \S\ref{sec:discussion}, as well as an alternative hypothesis for the origin of the Aquarius stream. In \S\ref{sec:conclusion} we present our conclusions and critical interpretations.

@@

\subsection{Radial Velocities}
\label{sec:radial-velocities}
The radial velocity for each star was determined in a two-step method. An initial estimate of the radial velocity was ascertained by cross-correlation with a synthetic spectrum of a giant with $T_{eff} = 4500$\,K, $\log{g} = 1.5$, and [M/H] $= -1.0$ across the wavelength range $8450 - 8700$ \AA{}, allowing us to place our spectrum at pseudo-rest. Velocities found from cross-correlation are typically within 1\,km\,s$^{-1}$ of our final published value.

The equivalent widths for $\sim{}N$ absorption lines were measured, allowing for the radial velocity of each line to be calculated from the ratio of laboratory and rest wavelengths. The final radial velocity is then given by the mean of these $\sim{}N$ measurements. Figure \ref{fig:line-velocities} shows the line velocities for HD440077 after  being placed at rest using our cross-correlation velocity of X.XX\,km s$^{-1}$. As expected, the mean offset is small ($-1.1$\,km s$^{-1}$), and our standard deviation is X.XX\,km s$^{-1}$ from N line measurements. This provides us with a final measured radial velocity of $X.XX \pm X.XX$\,km s$^{-1}$. This process was repeated for all standards and observations. The radial velocities published in Table \ref{tab:observations} are the final values from this two-step method. Heliocentric velocities agree excellently () with previously reported literature values \citep{Williams;wylie-de_boer;2012}.

@@

% equivalent widths plot with Norris for HD 122563

% equivalent widths table

\subsection{Model Atmospheres}
We employed the ATLAS9 stellar atmospheres of \citet{Castelli;Kurucz_2003} for this analysis. These one-dimensional models ignore any centre-to-limb spatial variations, assume local thermodynamic equilibrium as well as overall hydrostatic equilibrium, and assumes no convective overshoot from the photosphere. The stellar parameter spacing for these grid models is 250\,K, 0.5 dex in surface gravity, 0.5 dex in [M/H] and 0.1 dex in [$\alpha$/Fe]. We interpolated the stellar densities, temperatures, electron pressures at all stellar depths between atmospheres using the Quickhull algorithm during our analyses. Quickhull is reliant on Delaunay tessellation, which suffers from extremely skewed cells when the grid points vary in size by orders of magnitude -- as $T_{eff}$ values do compared to $\log{g}$ or [M/H]. If unaccounted for, these asymmetric cells can manifest as non-negligible errors in interpolated densities, temperatures, pressures, and opacities across all photospheric depths. We scaled each stellar parameter between zero and unity prior to interpolation to minimise these errors. During this step, $\log{g}$, [Fe/H] and [$\alpha$/Fe] points were scaled such that we simultaneously interpolate linearly in $T_{eff}$ and logarithmically in all other parameters.

\subsection{Stellar Parameters}
The most recent version of the spectral synthesis code MOOG \citep{Sneden;et-al_1973} has been used to derive individual line abundances and stellar parameters. This version employs a more accurate treatment of Rayleigh scattering \citet{Sobeck;et-al_2011}, which is particularly important for transitions blueward of 450\,nm. This is noteworthy, but is less relevant for this analysis as most of our line measurements blueward of 450\,nm are too blended and not used. We have scaled our abundances to solar values using the abundances of \citet{Asplund;et-al_2009}.

The effective temperature for each star was found by demanding a zero-trend in excitation potential and abundance for each measured Fe\,I line. In the same way, the microturbulence was found by demanding a zero trend in reduced equivalent width against abundance. Linear relationships with slopes of $|0.005|$ dex were considered to be converged. Surface gravity was found by forcing the mean Fe\,I and Fe\,II abundances to be equal, whilst maintaining zero trends with the excitation potential, reduced equivalent width and abundance. After these parameters had converged, the model metallicity was exactly matched to that of our line abundances. Line abundances that were unusually deviant ($>3\sigma$) from the mean were removed. The largest number of outlier measurements removed for any observation was N for X.

% spectral plots for five stars

\begin{deluxetable*}{lcccccccccccccc}
\tablecolumns{2}
\tabletypesize{\scriptsize}
\tablecaption{Stellar Parameters}
\tablehead{
	\colhead{Designation} &
	\colhead{$T_{\mbox{eff}}$} &
	\colhead{$\log{g}$} &
	\colhead{$v_t$} &
	\colhead{[Fe/H]} & 
	\colhead{$T_{\mbox{eff}}$} &
	\colhead{$\log{g}$} &
	\colhead{$v_t$} &
	\colhead{[Fe/H]} &
	\colhead{Source} \\
}
\startdata
C2225316-14437	& 4350	& 1.60	& 1.80	& --1.19	
				& 4235 $\pm$ 118 & 1.45 $\pm$ 0.21 & 1.96 $\pm$ 0.11 & --1.20 $\pm$ 0.14 \\
C2306265-085103	& 4180	& 1.00	& 1.80 	& --1.16
				& \dots	& \dots	& \dots	& \dots
				& \dots \\	
HD41667			& 4630	& 1.70	& 1.66 	& --1.21
				& 4605	& 1.88	& 1.44	& --1.16
				& \citet{Gratton;et-al_2000} \\
HD44007			& 4790	& 1.78	& 1.63	& --1.80
				& 4850	& 2.00 	& 2.20	& --1.71
				& \citet{Fulbright_2000} \\
HD142948		& 4950	& 2.19	& 1.78	& --0.79
				& 4713 	& 2.17 	& 1.38	& --0.77
				& \citet{Gratton;et-al_2000} \\
J221821-183424	& 4570	& 0.80	& 1.93	& --1.60
				& 4395 $\pm$ 205 & 1.45 $\pm$ 0.35 & 1.96 $\pm$ 0.18 & --1.15 $\pm$ 0.21
				& \citet{wylie-de-boer;et-al_2012} \\
J223504-152834	& 4620	& 2.15 	& 1.44	& --0.65
				& 4597 $\pm$ 158 & 2.40 $\pm$ 0.14 & 1.47 $\pm$ 0.07 & --0.98 $\pm$ 0.17
				& \citet{wylie-de-boer;et-al_2012} \\
J223811-104126	& 5100	& 3.00	& 1.21	& --1.44
				& 5646 $\pm$ 147 & 4.60 $\pm$ 0.15 & 1.09 $\pm$ 0.11 & --1.20 $\pm$ 0.20
				& \citet{wylie-de-boer;et-al_2012} 
\enddata

\end{deluxetable*}

@@

% elemental abundances

% comparison to literature for a given star

% chemical synthesis

\subsection{Distances}
% isochrone fitting

% discuss Williams et al distances

\subsection{Dynamics}

% U, V, W calculations

@@

% wylie de boer discrepancies

@@

% Chemistries for various GC trends

% chemistries - [Ni/Na]

% chemistries - [Al/Mg]

% which GC's do not show these trends

% closest looking globular clusters from Harris catalogue

% could they be displaced thick disk stars?
